---
title: Impact of Irma on Insurer Stock Prices - Part 1
author: ''
date: '2017-09-08'
slug: impact-of-irma-on-insurer-stock-prices-part-1
categories: []
tags:
  - stocks
subtitle: ''
draft: false
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)

library(tidyverse)
library(tidyquant)
library(timetk)
library(CausalImpact)
library(DT)
library(widgetframe)
library(kableExtra)

```

P&C (re)insurer stock price reactions to Hurricane Irma. Below we show the cumulative change in stock price from September 4 thru September 8. 

```{r set-dates}

start_event <- as_date("2017-09-05")
end_event <- as_date("2017-09-09")
start_stock <- as_date("2017-01-01")
end_stock <- as_date("2017-08-22")
start_graph <- end_event - weeks(6)

```


```{r import-tickers}

# get a list of all the Property-Casualty Insurers from the major US stock exchanges
stock_tickers <- list('AMEX','NASDAQ','NYSE') %>% map_df(tq_exchange) %>% 
  filter((industry == 'Property-Casualty Insurers' | symbol %in% c("AGII", "ERIE", "AIZ")) & !is.na(market.cap)) %>%
  mutate(market_cap = if_else(stringr::str_sub(market.cap,start = -1)=="M",1e6,1e9) *
           as.numeric(stringr::str_sub(market.cap,start = 2, end = -2))) %>%
  dplyr::select(symbol, company, sector, industry, market_cap) %>%
  arrange(desc(market_cap))

# get shares outstanding from most recent 10Q balance sheet; per Google finance
stock_shares <- stock_tickers %>% tq_get(get = "financials") %>% group_by(symbol) %>% 
  filter(type == "BS") %>% mutate(shares = map_dbl(quarter, ~ .x$value[.x$group == 42][1] * 1e6)) %>%
  dplyr::select(symbol, shares)

# not able to get XL shares outstanding, so manually put it in as of Q2 2017
stock_tickers <- stock_tickers %>% left_join(stock_shares, by = "symbol") %>%
  mutate(shares = if_else(symbol == "XL", 257.94 * 1e6, shares)) %>% filter(!is.na(shares))

```


```{r import-stock-prices}

# now get the historical stock prices for all of the tickers; calculate daily returns
stock_prices <- stock_tickers %>% tq_get(get = "stock.prices", from = start_stock, to = end_event) %>% 
  filter(!is.na(adjusted)) 

stock_prices <- stock_prices %>% timetk::tk_tbl() %>% dplyr::group_by(symbol) %>%
  tq_mutate(select = adjusted, mutate_fun = periodReturn, period = "daily", type = "arithmetic", 
            col_rename = "daily_returns")

# get S&P 500 historical prices; use for comparison and CausalImpact calculation
sp500_price <- tibble(symbol = "SPY") %>% tq_get(get = "stock.prices", 
                                                 from = start_stock, to = end_event) %>%
  tq_mutate(select = adjusted, mutate_fun = periodReturn, period = "daily", type = "arithmetic", 
            col_rename = "daily_returns")

sp500_price <- sp500_price %>% mutate(sp500 = adjusted, sp500_returns = daily_returns)
stock_prices <- stock_prices %>% left_join(sp500_price %>% dplyr::select(date, sp500, sp500_returns), by = "date")

```


```{r cum-change}

stock_prices <- stock_prices %>% mutate(post_event = if_else(date >= start_event, 1, 0)) %>%
  group_by(symbol, post_event) %>% mutate(cum_return = if_else(post_event == 1, cumprod(1 + daily_returns) - 1, 0)) %>%
  mutate(sp500_cum_return = if_else(post_event == 1, cumprod(1 + sp500_returns) - 1, 0)) %>% ungroup()

impact_of_event <- stock_prices %>% filter(post_event == 1) %>% group_by(symbol) %>%
  summarise(company = company[1], start_price = first(open), end_price = last(adjusted), 
            cum_return = last(cum_return), sp500_cum_return = last(sp500_cum_return),
            market_cap_change = shares[1] * first(open) * cum_return, market_cap = market_cap[1]) %>% ungroup() %>% 
  arrange(cum_return)

```


# Summary table

The table below shows the cumulative impact on stock prices for all U.S. P&C (re)insurers. The final column includes the change in market capitalization.

#### Impact of Hurricane Irma on P&C (Re)Insurer Stock Prices

```{r summary-datatable}

dt_impact_of_event <- impact_of_event %>% dplyr::select(-sp500_cum_return, - market_cap) %>% 
  mutate(market_cap_change = market_cap_change / 1e6) %>%
  DT::datatable(  
  options = list(
     columnDefs = list(list(className = 'dt-left')),
     pageLength = 15, lengthMenu = c(10, 15, 20, 25)),
  fillContainer = TRUE, rownames = FALSE,
  caption = htmltools::tags$caption(
    style = 'caption-side: top; text-align: left;', htmltools::em('from September 4, 2017 to September 7, 2017')),
  colnames = c("Ticker", "Company", "Initial stock price", "End stock price", "% change", "Market cap change ($m)")) %>%
    formatCurrency(c(3:4), '', digits = 2) %>% formatCurrency(6, '', digits = 1) %>% formatPercentage(5, digits = 1)

frameWidget(dt_impact_of_event, width = '100%')

```

<br>

```{r summary-graph}

impact_of_event %>% mutate(symbol = factor(symbol, levels = rev(impact_of_event$symbol))) %>%
  slice(1:30) %>%
  ggplot(aes(y = cum_return, x = symbol)) +
  geom_col(fill = "#0083A9") + geom_text(aes(label = scales::percent(round(cum_return,3))), size = 3, nudge_y = -0.01) +
  coord_flip() + AonECM::ggThemeAon() +
  scale_y_continuous(labels = scales::percent_format()) + xlab(NULL) + ylab("Cumulative stock price change since event") +
  ggtitle("Largest cumulative stock price changes", subtitle = "due to Hurricane Irma") +
  labs(caption = "*from September 4, 2017 to September 7, 2017")

```


```{r total-value-lost}

total_value_lost <- -sum(impact_of_event$market_cap_change[impact_of_event$market_cap_change<0])
total_value_lost <- paste0(scales::dollar(total_value_lost/1e9), "B")

```

This analysis highlights the impact that catastrophe losses - in this case still perceived losses - can have on P&C (re)insurer stock prices. The estimated total market value lost is `r total_value_lost`.

