---
title: "Impact of Earnings Volatility on Insurer Valuations - Part 3"
subtitle: "Varying parameters by earnings volatility"
date: 2018-03-16
slug: impact-of-earnings-volatility-on-insurer-valuations-part-3
tags:
    - insurance
    - stocks
draft: false
---

```{r setup, warning=FALSE,message=FALSE,include=FALSE}

library(tidyverse)
library(readxl)
library(AonInternal)
library(lme4)
library(rstanarm)
library(flextable)
library(AonECM)
# library(officer)
# library(rvg)
library(broom)
library(stringr)
library(rlang)

SetTheme()
AonECM::ggSetTheme()

knitr::opts_chunk$set(
	fig.align = "center",
	message = FALSE,
	warning = FALSE,
	echo = FALSE
)

p_to_b_format <- function() {
    function(x) {
        if (length(x) == 0) 
            return(character())
        x <- round(x, 2)
        paste0(x, "x")
    }
}


```

# Goal

Prior posts in this series: [Part 1](`r blogdown::shortcode_html("relref", "2018-02-11-impact-of-earnings-volatility-on-insurer-valuations.html")`), [Part 2](`r blogdown::shortcode_html("relref", "2018-03-15-impact-of-earnings-volatility-on-insurer-valuations-part-2.html")`)

In this post, we will test whether using earnings coefficient of variation (CV) as a categorical variable in our regression improves the model performance. Specifically, we will split earnings CVs by company into buckets and build a hierarchical regression model with slope and intercept that vary by earnings CV bucket.

# Data

The table below summarizes the data required for this analysis, which is the same as used in the prior posts. This table shows the data for the "P&C Commercial" lines sector, but global data across 10 P&C sectors are used in the analysis.

```{r import-data}

data <- readxl::read_xlsx("data/ABS_Weekly_data.xlsx", sheet = "ABS_weekly", range = "H3:AG157") %>%
  select(-c(8:11)) %>%
  mutate_at(-c(1:2), as.double)

data <- data %>%
  filter(!(sector %in% c("FMG","Health","LargeCap", "Title"))) %>%
  filter(include == 1) %>%
  mutate(spread = roae_2018 - cost_of_equity) %>%
  drop_na(p_to_b, roae_2018, eps_vol_5, p_to_tbv)

sector_summary <- data %>% 
  group_by(sector) %>% 
  summarise(n_comps = n(), median_pb = median(p_to_b, na.rm = TRUE), median_ptbv = median(p_to_tbv, na.rm = TRUE),
            median_roae = median(roae_2018, na.rm = TRUE), median_cost_equity = median(cost_of_equity, na.rm = TRUE),
            median_spread = median(spread, na.rm = TRUE),
            mean_eps_vol = mean(eps_vol_5, na.rm = TRUE), median_eps_vol = median(eps_vol_5, na.rm = TRUE))

data <- data %>% 
  left_join(sector_summary, by = "sector") %>%
  mutate(high_vol2 = if_else(eps_vol_5 > median_eps_vol, 1, 0))

# add eps vol quintile column
data <- data %>% 
  mutate(pb_ntile = ntile(p_to_b, 4),
         vol_ntile = ntile(eps_vol_5, 4))

```


```{r pb-vs-roe-table}

data_ft <- data %>%
  # filter(sector %in% c("APAC") & p_to_b < 6) %>%
  drop_na(roae_2018, eps_vol_5) %>%
  arrange(sector, desc(market_cap)) %>%
  transmute(Sector = sector, Company = company, 
         `Market Cap ($b)` = AonECM::number_format(market_cap/1000,1), 
         `P:B Ratio` = paste0(p_to_b, "x"), 
         `P:TBV Ratio` = paste0(p_to_tbv, "x"), 
         `2018 ROAE` = AonECM::pct_format(roae_2018/100, 1),
         `20 qrt EPS CV` = AonECM::pct_format(eps_vol_5/100, 1))

ft_pb_vs_roae1 <- data_ft %>% 
  filter(Sector == "P&C_Comm") %>%
  flextable::regulartable() %>%
  AonECM::ft_theme_aon() %>%
  flextable::align(j = 1:2, part = "all") %>%
  AonECM::add_title_header("By company input data - P&C commercial lines")

ft_pb_vs_roae1


```
<br>

The data elements include: current P:B ratio, 2018 prospective ROE, and historical EPS CV over the past 20 quarters. The data used in this analysis is as of December 15, 2017.

The table below summarizes the input data by price-to-book value quartiles. It highlights the relationship between P:B ratio, ROAE, and EPS CV. As the P:B ratio increases, ROAE increases and EPS CV decreases. These are the effects that we want to capture in our model.

```{r eps-vol-by-sector}

summary_ntiles <- data %>%
  group_by(pb_ntile) %>%
  summarise(`# of companies` = n(), 
            `Median P:B Ratio` = paste0(round(median(p_to_b),1), "x"), 
            `Median P:TBV Ratio` = paste0(round(median(p_to_tbv),1), "x"),
            `Median 2018 ROAE` = AonECM::pct_format(median(roae_2018)/100, 1),
            `Median EPS CV` = AonECM::pct_format(median(eps_vol_5)/100, 1))

ft_ntiles <- summary_ntiles %>% 
  flextable::regulartable() %>%
  AonECM::ft_theme_aon() %>%
  flextable::align(j = 1, part = "all") %>%
  AonECM::add_title_header("Summary of metrics by P:B quartiles")

ft_ntiles

```
<br>

# Current model

THe best model from the prior posts used ROE and an indicator variable for EPS volatility. A company was deemed "high" volatility if their 20 quarter EPS CV from greater than the sector median.  

```{r fit-stan-pb-eps, include=FALSE}

seed <- 12

fit_pb_roae_eps <- stan_glmer(p_to_b ~ 1 + (1 + roae_2018 + roae_2018:high_vol2 | sector),
                          family = Gamma(link = "identity"),
                          prior = student_t(df = 3, location = 0, scale = 2.5),
                          prior_intercept = student_t(df = 3, location = 0, scale = 2.5),
                          adapt_delta = 0.999,
                          data = data,
                          seed = seed)

```

```{r fit-stan-pb-vol, include=FALSE}

fit_pb_roae_vol <- stan_glmer(p_to_b ~ 1 + (1 + roae_2018 | sector) + (0 + roae_2018 | vol_ntile),
                          family = Gamma(link = "identity"),
                          prior = student_t(df = 3, location = 0, scale = 2.5),
                          prior_intercept = student_t(df = 3, location = 0, scale = 2.5),
                          adapt_delta = 0.99,
                          data = data,
                          seed = seed)

```

```{r param-summary-fns}

add_bayes_R2 <- function(group_var, stan_model){

  group_var_enq <- enquo(group_var)

  data <- stan_model$data
  vars <- all.vars(terms(stan_model$formula))
  col_name <- rlang::sym(vars[match("sector", vars)])
  
  new_data <- data %>% 
    filter((!!col_name) == (!!group_var_enq)) %>%
    select(!!vars)
  
  stan_model %>% 
    bayes_R2(newdata = new_data) %>%
    median(.)
  
}

add_bayes_R22 <- function(sector, ntile, stan_model){

  sector_enq <- enquo(sector)
  ntile_enq <- enquo(ntile)
  
  data <- stan_model$data
  vars <- all.vars(terms(stan_model$formula))

  new_data <- data %>% 
    filter(sector == (!!sector_enq) & vol_ntile == (!!ntile_enq)) %>%
    select(!!vars)
  
  stan_model %>% 
    bayes_R2(newdata = new_data) %>%
    median(.)
  
}

add_sector_ppd_sims <- function(sector, stan_model, draws = 10){
  
  vars <- all.vars(terms(stan_model$formula))
  new_roae <- seq(from = 2, to = 34, by = 2)
  new_data <- tibble(sector = sector, roae_2018 = new_roae)
  
  if(length(vars) > 3){
    new_data <- new_data %>% mutate(high_vol2 = 0)
    new_data2 <- new_data %>% mutate(high_vol2 = 1)
    new_data <- new_data %>% bind_rows(new_data2)
  }
  
  stan_model %>% 
    posterior_predict(newdata = new_data, draws = draws) %>%
    as_tibble() %>%
    rowid_to_column(var = "trial") %>%
    gather(id, fitted, -trial, convert = TRUE) %>%
    left_join(new_data %>% mutate(id = 1:nrow(.)), by = "id") %>%
    drop_na(fitted)
  
}

add_sector_ppd_sims2 <- function(sector, ntile, stan_model, draws = 10){

  vars <- all.vars(terms(stan_model$formula))

  new_roae <- seq(from = 2, to = 34, by = 2)
  new_data <- tibble(sector = sector, vol_ntile = ntile, roae_2018 = new_roae)

  stan_model %>% 
    posterior_predict(newdata = new_data, draws = draws) %>%
    as_tibble() %>%
    rowid_to_column(var = "trial") %>%
    gather(id, fitted, -trial, convert = TRUE) %>%
    left_join(new_data %>% mutate(id = 1:nrow(.)), by = "id") %>%
    drop_na(fitted)
  
}

```

```{r create-param-summaries}

pb_roae_eps_params <- coef(fit_pb_roae_eps)$sector %>% 
  as_tibble() %>% 
  rownames_to_column(var = "sector") %>%
  rename(intercept = `(Intercept)`, roae_high_vol = `roae_2018:high_vol2`) %>%
  mutate(bayes_R2 = map_dbl(sector, add_bayes_R2, fit_pb_roae_eps)) %>%
  mutate(ppd_sims = map(sector, add_sector_ppd_sims, fit_pb_roae_eps, draws = 1000)) %>%
  mutate(ppd_summary = map(ppd_sims, ~ .x %>% group_by(id) %>% 
                             summarise(n_points = length(id),
                                       roe = roae_2018[1],
                                       high_vol = high_vol2[1],
                                       median = median(fitted),
                                       lower = quantile(fitted, 0.1),
                                       upper = quantile(fitted, 0.9))))

pb_roae_vol_params <- data %>% 
  distinct(sector, vol_ntile) %>% 
  left_join(coef(fit_pb_roae_vol)$sector %>% 
              as_tibble() %>% 
              rownames_to_column(var = "sector") %>%
              rename(intercept = `(Intercept)`), by = "sector") %>%
  left_join(coef(fit_pb_roae_vol)$vol_ntile %>% 
              as_tibble() %>% 
              rownames_to_column(var = "vol_ntile") %>%
              mutate(vol_ntile = as.integer(vol_ntile)) %>%
              select(vol_ntile, roae_vol = roae_2018), by = "vol_ntile") %>%
  mutate(bayes_R2 = map2_dbl(sector, vol_ntile, add_bayes_R22, fit_pb_roae_vol)) %>%
  mutate(ppd_sims = map2(sector, vol_ntile, add_sector_ppd_sims2, fit_pb_roae_vol, draws = 1000)) %>%
  mutate(ppd_summary = map(ppd_sims, ~ .x %>% group_by(id) %>% 
                             summarise(n_points = length(id),
                                       roe = roae_2018[1],
                                       median = median(fitted),
                                       lower = quantile(fitted, 0.1),
                                       upper = quantile(fitted, 0.9))))

```


```{r create-ppd-fn, include=FALSE}

create_ppd_sims <- function(stan_model, draws = 10, variable, actual){
  
  data <- stan_model$data
  
  ppd_sims <- stan_model %>% posterior_predict(draws = draws) %>% as_tibble() %>%
  rowid_to_column(var = "trial") %>%
  gather(company_id, fitted, -trial, convert = TRUE)
  
  ppd_sims$company <- data$company[ppd_sims$company_id]
  ppd_sims$sector <- data$sector[ppd_sims$company_id]
  ppd_sims <- ppd_sims %>%
    left_join(data %>% select_("sector", "company", variable = variable, actual = actual, "high_vol2", "vol_ntile"), 
              by = c("company", "sector"))

  ppd_summary <- ppd_sims %>% 
    group_by(company_id, variable) %>% 
    summarise(company = company[1], sector = sector[1], actual = actual[1],
              high_vol = as.factor(high_vol2[1]),
              vol_ntile = as.factor(vol_ntile[1]),
              median_fitted = median(fitted), 
              pctl_25 = quantile(fitted, 0.1), pctl_75 = quantile(fitted, 0.9))
  
  list(ppd_sims = ppd_sims,
       ppd_summary = ppd_summary %>% ungroup())
  
}


```

```{r examine-stan-pb}

ppd_pb_roae_eps <- create_ppd_sims(fit_pb_roae_eps, 4000, "roae_2018","p_to_b")
ppd_pb_roae_vol <- create_ppd_sims(fit_pb_roae_vol, 4000, "roae_2018","p_to_b")

```

Add graph for model with earnings volatility indicator variable. 

```{r gg-baseline}

gg_ppc_roae_eps <- ppd_pb_roae_eps$ppd_summary %>% 
  ggplot(aes(y = actual, x = variable / 100)) + 
  # geom_ribbon(aes(y = median, ymin = lower, ymax = upper, x = roe/100, group = high_vol, fill = high_vol), 
  #             alpha = 0.25,
  #             data = pb_roae_eps_params %>% unnest(ppd_summary) %>% mutate(upper = pmin(upper, 5))) +
  geom_abline(aes(intercept = intercept, slope = roae_2018*100),
              data = pb_roae_eps_params, colour = aon.colors$teal) +
  geom_abline(aes(intercept = intercept, slope = (roae_2018+roae_high_vol)*100),
              data = pb_roae_eps_params, colour = aon.colors$lightgray) +
  geom_point(aes(group = high_vol, colour = high_vol)) + 
  facet_wrap(~ sector) +
  scale_color_manual(name = NULL, 
                     labels = c("Low volatility", "High volatility"),
                     values = c(aon.colors$teal, aon.colors$lightgray)) +
  scale_x_continuous(labels = scales::percent_format(), limits = c(0.025,0.25)) + 
  scale_y_continuous(labels = p_to_b_format(), limits = c(0,5)) + 
  xlab("2018 Prospective Return on Average Equity") + ylab("Price to Book Ratio") +
  theme(plot.title = element_text(colour = aon.colors$teal), plot.subtitle = element_text(colour = aon.colors$teal)) +
  labs(title = "Fitted model: P:B Ratio and Prospective ROE by Sector", 
       subtitle = "With a variable for EPS volatility")

gg_ppc_roae_eps

gg_ppc_roae2 <- ppd_pb_roae_vol$ppd_summary %>% 
  ggplot(aes(y = actual, x = variable / 100)) + 
  # geom_ribbon(aes(y = median, ymin = lower, ymax = upper, x = roe/100), 
  #             fill = aon.colors$lightgray, alpha = 0.25, 
  #             data = pb_roae_vol_params %>% unnest(ppd_summary) %>% mutate(upper = pmin(upper, 5))) +
  geom_abline(aes(intercept = intercept, slope = (roae_2018+roae_vol)*100, group = as.factor(vol_ntile), colour = as.factor(vol_ntile)),
              data = pb_roae_vol_params) +
  geom_point(aes(group = vol_ntile, colour = vol_ntile)) + 
  facet_wrap(~ sector) +
  scale_x_continuous(labels = scales::percent_format(), limits = c(0.025,0.25)) + 
  scale_y_continuous(limits = c(0,5)) + 
  xlab("2018 Prospective Return on Average Equity") + ylab("Price to Tangible Book Ratio") +
  ggtitle("Fitted model: P:BV Ratio and Prospective ROE by Earnings CV")

gg_ppc_roae2

ggplot() + 
  geom_abline(aes(intercept = intercept, slope = (roae_2018+roae_vol)*100, group = as.factor(vol_ntile), colour = as.factor(vol_ntile)),
              data = pb_roae_vol_params) +
  facet_wrap(~ sector)
  
```

# Alternative model

For the alternative model, we use the 20 quarter EPS CV quartile for each company.

Add graph with lines for each volatility level.
Add table with impact on pb by sector at 10% ROE and the 4 earnings vol quartiles.
State that loo indicates a better fitting model.

```{r compare-bayes-r2}

pb_roae_params %>%
  left_join(pb_roae_vol_params %>% select(sector, bayes_R2_2 = bayes_R2), by = "sector") %>%
  transmute(Sector = sector,
            `Baseline` = AonECM::pct_format(bayes_R2,0),
            `P:TBV` = AonECM::pct_format(bayes_R2_2,0),
            `P:TBV vs. Spread` = AonECM::pct_format(bayes_R2_3,0)) %>%
  flextable::regulartable() %>%
  AonECM::ft_theme_aon() %>%
  flextable::align(j = 1, part = "all") %>%
  AonECM::add_title_header("Comparison of Bayesian R2 by model")

```
<br>

# Conclusion


