---
title: Insurer Stock Buyback ROI
subtitle: ''
date: '2018-03-07'
slug: insurer-stock-buyback-roi
tags:
  - insurance
  - stocks
draft: yes
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}

library(knitr)
library(AonInternal)
library(tidyverse)
library(readxl)
library(tibbletime)
library(lubridate)
library(tidyquant)

knitr::opts_chunk$set(
	fig.align = "center",
	message = FALSE,
	warning = FALSE,
	echo = FALSE
)

AonInternal::SetTheme()
aon_colors <- c("#0083A9", "#808080", "#BFBFBF", "#5EB6E4", "#F0AB00", "#7AB800", 
                "#6E267B", "#BC5FCD","#E9CAEE", "#D3CD8B", "#7E7830", "#3E5B00", "#E11B22")

AonECM::ggSetTheme()

```

# Goal

Calculate historical return on share buybacks for P&C (re)insurance companies. 

- Explain methodology by walking through data and calculations for a company
- Apply methodology to all stocks and show results

# Data used

- Historical buybacks, common shares outstanding, common div, special dividends, share price

```{r import-data}

data <- readxl::read_xlsx("data/insurer_buybacks.xlsx", sheet = "import") %>%
  gather(key = "statistic", value = "value", -ticker, -company) %>%
  separate(statistic, c("yr_qtr","metric"), sep = "_") %>%
  separate(yr_qtr, c("year","qtr"), sep = 4, remove = FALSE) %>%
  mutate(value = as.double(value), 
         year = as.double(year),
         date = lubridate::yq(yr_qtr), 
         qtr_ = lubridate::quarter(date)) %>%
  replace_na(list(value = 0)) %>% 
  spread(key = metric, value = value) %>%
  mutate(dollarrepurchase = avgpricepershare * sharesrepurchased)

```

```{r import-stock-prices}

# get a list of all the Property-Casualty Insurers from the major US stock exchanges
stock_tickers <- list('AMEX','NASDAQ','NYSE') %>% map_df(tq_exchange) %>% 
  filter((industry == 'Property-Casualty Insurers' | symbol %in% c("AGII", "ERIE", "AIZ")) & !is.na(market.cap)) %>%
  mutate(market_cap = if_else(stringr::str_sub(market.cap,start = -1)=="M",1e6,1e9) *
           as.numeric(stringr::str_sub(market.cap,start = 2, end = -2))) %>%
  dplyr::select(symbol, company, sector, industry, market_cap) %>%
  arrange(desc(market_cap))

stock_tickers <- stock_tickers %>% 
  inner_join(data %>% distinct(ticker), by = c("symbol"="ticker"))

# now get the historical stock prices for all of the tickers; get quarter close stock prices
stock_prices <- stock_tickers %>%
  tq_get(get = "stock.prices", from = "2010-01-01", to = "2017-12-31") %>% 
  filter(!is.na(adjusted)) %>% 
  select(symbol, company, market_cap, date, close, adjusted) %>%
  group_by(symbol) %>%
  as_tbl_time(index = date) %>%
  as_period("quarterly", side = "end") %>%
  mutate(year = lubridate::year(date), qtr_ = lubridate::quarter(date))

```

```{r combine-date}

data_to_use <- data %>% 
  select(-date) %>% 
  left_join(stock_prices %>% 
              select(symbol, market_cap, year, date, qtr_, close, adjusted), 
            by = c("ticker"="symbol", "year", "qtr_")) %>%
  drop_na() %>%
  group_by(ticker) %>%
  as_tbl_time(index = date) %>%
  filter_time('2012' ~ '2016')

data_to_use <- data_to_use %>%
  mutate(cum_shares_repurchased = cumsum(sharesrepurchased)) %>%
  mutate(cash_flow = -dollarrepurchase + (commondivpershare + specdivpershare) * cum_shares_repurchased,
         cash_flow_final = cash_flow + if_else(year == 2016 & qtr_ == 4, close * cum_shares_repurchased, 0),
         cash_flow_final2 = cash_flow + if_else(year == 2016 & qtr_ == 4, adjusted * cum_shares_repurchased, 0))

```

# Methodology

Walk through methodology using TRV as an example. Use graphs to show dollar repurchased, shares repurchased, net cash flows.

Create for every company...nice single slide summary.

Show graph of stock price with avg repurchase price as a line; number of shares repurchased below it. Ideally include pb ratio graph also...

Bar graphs with quarterly cash flows - stacked to see impact of repurchase + dividends

```{r company-example}

company_example <- data_to_use %>%
  filter(ticker == "ACGL")

company_example %>%
  ggplot(aes(x = date, y = commonshares)) +
  geom_col()

company_example %>%
  ggplot(aes(x = date, y = sharesrepurchased)) +
  geom_col()

company_example %>%
  ggplot(aes(x = date, y = cum_shares_repurchased)) +
  geom_col()

company_example %>%
  ggplot(aes(x = date, y = cash_flow_final)) +
  geom_col()

company_example %>%
  ggplot(aes(x = date, y = cash_flow_final2)) +
  geom_col()


```

# Overall results

A few graphs or tables (datatable?) summarizing results for each company. Add any interesting commments.

```{r summary-table}

data_to_use %>% 
  summarise(n_qtrs = n(),
            market_cap = market_cap[1], init_shares = commonshares[1], 
            shares_repurchased = dplyr::last(cum_shares_repurchased), 
            dollar_repurchased = sum(dollarrepurchase), 
            pct_init_shares = shares_repurchased / init_shares, 
            pct_market_cap = sum(dollarrepurchase) / market_cap[1],
            tsr = ((sum((commondivpershare + specdivpershare) * commonshares) / dplyr::last(commonshares) + 
              dplyr::last(close)) / dplyr::first(close))^(1/(n_qtrs/4)) - 1,
            irr = (1+FinCal::irr(cash_flow_final))^4 - 1,
            buyback_effectiveness = (1+irr)/(1+tsr) - 1) %>%
  filter(pct_market_cap > 0.01) %>%
  arrange(desc(buyback_effectiveness))

```








