---
title: "Replicate Meyers' Changing Settlement Rate Backtest"
date: 2018-02-11
slug: replicate-meyers-changing-settlement-rate-backtest
tags:
    - insurance
    - reserving
draft: false
---

```{r setup, include=FALSE}

library(knitr)
library(AonInternal)
library(tidyverse)
library(reservetestr)
library(ChainLadder)
library(patchwork)

knitr::opts_chunk$set(
	fig.align = "center",
	message = FALSE,
	warning = FALSE,
	echo = FALSE
)

AonInternal::SetTheme()
aon_colors <- c("#0083A9", "#808080", "#BFBFBF", "#5EB6E4", "#F0AB00", "#7AB800", 
                "#6E267B", "#BC5FCD","#E9CAEE", "#D3CD8B", "#7E7830", "#3E5B00", "#E11B22")


```

[Meyers 2015]: http://www.casact.org/pubs/monographs/papers/01-Meyers.PDF
[Meyers 2016]: http://www.casact.org/pubs/forum/16wforum/Meyers.pdf
[Meyers 2016 spreadsheet]: http://www.casact.org/pubs/forum/16wforum/02b_Meyers_Dependencies_Appendix-10-13-2015.xls
[Meyers 2017 Actuarial Review]: http://ar.casact.org/actuarialreview/july_august_2017/MobilePagedArticle.action?articleId=1130425#articleId1130425

# Goal

The goal of this post is to replicate the univariate results from Glenn Meyers's paper [Dependencies in Stochastic Loss Reserve Models][Meyers 2016]. The stochastic reserving method used is called the Changing Settlement Ratio (CSR) model which was first described in [STOCHASTIC LOSS RESERVING USING BAYESIAN MCMC MODELS][Meyers 2015].

The R code provided with the [paper][Meyers 2016 spreadsheet] and from [Meyers' Actuarial Review Article][Meyers 2017 Actuarial Review] were used as references.

# Import data

Import the CAS loss reserve [database](http://www.casact.org/research/index.cfm?fa=loss_reserves_data) from the [`reservetestr`](https://github.com/problemofpoints/reservetestr) package.

Summary table shown below for full CAS database.

```{r import-data}

tri_data <- reservetestr::cas_loss_reserve_db

# summary of db
tri_data_summary <- tri_data %>% group_by(line) %>% 
  summarise(ct_comp = length(unique(company)), ct_acc_yr = length(unique(acc_yr)), ct_obs = n())

knitr::kable(tri_data_summary, format = "pandoc", digits = c(0,0,0,0),
             col.names = c("Line", "Number of companies", "Number of accident years", 
                           "Total number of observations"),
             format.args = list(big.mark = ","),
             caption = "Summary of CAS Loss Reserve Database")


```

Meyers only uses a subset of the database. He chose four lines of business - comauto, ppauto, wkcomp, othliab - and selected 50 companies within each line. We can filter the full database to include only these companies by using the spreadsheet provided on the [CAS website][Meyers 2016 spreadsheet], again pulling the data from the `reservetestr` package.

```{r filter-data}

tri_data_filt <- tri_data %>% get_meyers_subset()
  
# summary of db
tri_data_summary2 <- tri_data_filt %>% group_by(line) %>% 
  summarise(ct_comp = length(unique(company)), ct_acc_yr = length(unique(acc_yr)), ct_obs = n())

knitr::kable(tri_data_summary2, format = "pandoc", digits = c(0,0,0,0),
             col.names = c("Line", "Number of companies", "Number of accident years", 
                           "Total number of observations"),
             format.args = list(big.mark = ","),
             caption = "Summary of Data used in Meyers (2016)")


```

Now let's replicate Table 2.1 from [Meyers (2016)][Meyers 2016].

```{r table_2.1}

createTable2.1 <- function(tri_data2, comp_id, filter_line, use_line) {

  tri_data2 <- tri_data2 %>% filter(line == filter_line & group_id == comp_id)
  tri_format <- as.triangle(tri_data2 %>% 
              select(AY = acc_yr, dev = dev_lag, value = cum_paid_loss, net_ep, line), origin = "AY", lob = line)
  ay_names <- tibble(acc_yr = rownames(tri_format), 
                     premium = (tri_data2 %>% filter(dev_lag == 10) %>% 
                                  group_by(acc_yr) %>% summarise(premium = sum(net_ep)))$premium)
  tri_format <- as_tibble(tri_format) %>% bind_cols(ay_names) %>% select(acc_yr, premium, 1:10)
  
  return(
    
  knitr::kable(tri_format, format = 'pandoc', 
               format.args = list(big.mark = ","),
               caption = paste("Group", comp_id, "-", use_line, sep = " "))
  )
}

createTable2.1(tri_data_filt, 620, "comauto","CA")
createTable2.1(tri_data_filt, 620, "ppauto","PA")
createTable2.1(tri_data_filt, 1066, "comauto","CA")
createTable2.1(tri_data_filt, 1066, "ppauto","PA")

```

# The Changing Settlement Rate (CSR) Model

With our input data gathered, we can apply the CSR method to these 200 triangles. See the papers for details of the CSR method ([Dependencies in Stochastic Loss Reserve Models][Meyers 2016] and [STOCHASTIC LOSS RESERVING USING BAYESIAN MCMC MODELS][Meyers 2015]).

The number of iterations we ran was different from Meyers, so that increases the simulation error. We used `rstan`'s default assumptions of 4 chains, 2,000 iterations (1,000 warm up). 

# Model results comparison

The graph below compares our results with Meyers for Other Liability across four fitted parameters / model outputs. The results are reasonably close. The other lines look similar. 

```{r import-results}

csr_output <- readRDS("data/meyers_replicate_no_models.rds") %>%
  rename(group_id = id)
meyers_file <- reservetestr::meyers_2016_wintereforum_appendix %>%
  select(line = line2, group_id, mean = Estimate, sd = `Std Dev`, 
         outcome_pctl = Percentile, max_Rhat,
         delta = deltamean, gamma = gammamean) %>% 
  mutate(outcome_pctl = outcome_pctl/100)

```

```{r compare-results}

compare_to_meyers <- meyers_file %>% gather(stat, value_meyers, mean:gamma)
compare_to_meyers <- compare_to_meyers %>% 
  inner_join(csr_output %>% 
               select(-company, -data, -loo, -outcome, -div_base, -div_base2) %>% 
               gather(stat, value, mean:gamma), 
  by = c("line","group_id", "stat")) %>%
  mutate(value_meyers = if_else(stat %in% c("mean","sd"), value_meyers / 1e6, value_meyers)) %>%
  mutate(value = if_else(stat %in% c("mean","sd"), value / 1e6, value)) %>%
  filter(stat != "max_Rhat")

ggplot(data = filter(compare_to_meyers, line == 'OL')) + 
  geom_point(aes(x = value_meyers, y = value)) + 
  geom_abline(aes(intercept = 0, slope = 1), colour = aon.colors$darkgray) +
  facet_wrap(~ stat, scales = "free") +
  ylab("problemofpoints") + xlab("Meyers") +
  theme(plot.title = element_text(colour = aon.colors$teal), plot.subtitle = element_text(colour = aon.colors$teal)) +
  labs(title = "Comparison of key model results - Other Liability", 
       subtitle = "...results are in fairly good agreement", color = aon.colors$teal)


```

## Figure 3.1 Uniformity Tests for the CSR Model

Very consistent results shown below when compared to figure 3.1 from the paper.

```{r pp-plot}
# create pp plots 

csr_output <- csr_output %>% group_by(line) %>% arrange(outcome_pctl) %>%
  mutate(fitted = cumsum(rep(1/(n()+1), n()))) %>% arrange(line, group_id)
se_line <- csr_output %>% summarise(n = n(), se = 1.36 / sqrt(n))

csr_output <- csr_output %>% ungroup() %>% arrange(outcome_pctl) %>%
  mutate(fitted_all = cumsum(rep(1/(n()+1), n()))) %>% arrange(line, group_id)
se_line <- se_line %>% mutate(se_all = 1.36 / sqrt(nrow(csr_output)))

(ggplot(data = csr_output) 
 + geom_abline(colour=aon.colors$darkgray, size=0.65)
 + geom_abline(data=se_line, aes(intercept=se, slope=1), 
               colour=aon.colors$darkgray, size=0.65, linetype=5)
 + geom_abline(data=se_line, aes(intercept=-se, slope=1), 
               colour=aon.colors$darkgray, size=0.65, linetype=5)
 + geom_point(aes(x = fitted, y = outcome_pctl))
 + facet_wrap(~ line)
 + xlab("Expected Percentile") + ylab("Predicted Percentile")  #+ scale_y_continuous(labels=NULL) 
 + theme(plot.title = element_text(colour = aon.colors$teal), plot.subtitle = element_text(colour = aon.colors$teal))
 + ggtitle("PP Plot - Predicted vs. Expected Percentiles")
)

```

Same goes for the pp-plot and histogram with all the triangles included.

```{r pp-plot-all, fig.height=3.2, fig.width=7}

gg_pp_all <- (ggplot(data = csr_output) +
 geom_abline(colour=aon.colors$darkgray, size=0.65)
 + geom_abline(data=se_line, aes(intercept=se_all, slope=1), 
               colour=aon.colors$darkgray, size=0.65, linetype=5)
 + geom_abline(data=se_line, aes(intercept=-se_all, slope=1), 
               colour=aon.colors$darkgray, size=0.65, linetype=5)
 + geom_point(aes(x = fitted_all, y = outcome_pctl))
 # + facet_wrap(~ line)
 + xlab("Expected Percentile") + ylab("Predicted Percentile")  #+ scale_y_continuous(labels=NULL) 
 + theme(plot.title = element_text(colour = aon.colors$teal), plot.subtitle = element_text(colour = aon.colors$teal))
 + ggtitle("PP Plot and Histogram of Predicted Percentiles - All Lines")
)

# create histogram of predicted percentiles
gg_hist_all <- (ggplot(data = csr_output) +
 geom_histogram(aes(x = outcome_pctl), fill=aon.colors$lightgray, bins=12) 
 # + facet_wrap(~ line)
 + xlab("Predicted Percentile") + ylab(NULL)  #+ scale_y_continuous(labels=NULL) 
 + xlim(0, 1)
 # + theme(plot.title = element_text(colour = aon.colors$teal), plot.subtitle = element_text(colour = aon.colors$teal))
 # + ggtitle("Histogram of Predicted Percentiles - All lines")
)

gg_pp_all + gg_hist_all

```

## Figure 3.2 Standardized Residual Plots for the CSR Model

```{r stdresid, fig.height=6.5, fig.width=8.0}

## ---- resid_ggplot ----
ggResidPlot <- function(df, x_var, x_lab = NULL, y_lab = NULL){
  
  gg <- ggplot(data = df, aes_string(y = "stdresid")) +
    geom_hline(aes(yintercept = 0), colour=aon.colors$darkgray, size=0.65) +
    geom_hline(aes(yintercept = -3), colour=aon.colors$darkgray, size=0.65) +
    geom_hline(aes(yintercept = 3), colour=aon.colors$darkgray, size=0.65) +
    geom_point(aes_string(x = x_var)) +
    facet_wrap(~ label) +
    ylim(-4.0, 4.0) + scale_x_continuous(breaks = c(2,4,6,8,10)) +
    xlab(x_lab) + ylab(y_lab)
    # theme(plot.title = element_text(colour = aon.colors$teal), plot.subtitle = element_text(colour = aon.colors$teal)) +
    # ggtitle(paste0(x_lab," - standardized residual plots"))
  
  gg
  
}

## calculate standardized residuals ####
n_samples <- 100
data_resid_plot <- readRDS("data/meyers_replicate_resid_plots.rds") %>%
  mutate(label = paste0(line, " ", id))

# create std residual plots
ggResidPlot(data_resid_plot %>% filter(line == "CA" & id == 620), "w", NULL) +
ggResidPlot(data_resid_plot %>% filter(line == "CA" & id == 620), "d", NULL) +
ggResidPlot(data_resid_plot %>% filter(line == "CA" & id == 620), "cy", NULL) +
ggResidPlot(data_resid_plot %>% filter(line == "PA" & id == 620), "w", NULL) +
ggResidPlot(data_resid_plot %>% filter(line == "PA" & id == 620), "d", NULL) +
ggResidPlot(data_resid_plot %>% filter(line == "PA" & id == 620), "cy", NULL) +
ggResidPlot(data_resid_plot %>% filter(line == "CA" & id == 1066), "w", NULL,"Standardized residuals") +
ggResidPlot(data_resid_plot %>% filter(line == "CA" & id == 1066), "d", NULL) +
ggResidPlot(data_resid_plot %>% filter(line == "CA" & id == 1066), "cy", NULL) +
ggResidPlot(data_resid_plot %>% filter(line == "PA" & id == 1066), "w", "Accident year") +
ggResidPlot(data_resid_plot %>% filter(line == "PA" & id == 1066), "d", "Development year") +
ggResidPlot(data_resid_plot %>% filter(line == "PA" & id == 1066), "cy", "Calendar year") +
  plot_layout(nrow = 4)
  
```

All the standardized residuals plots appear consistent with the paper.

# Conclusion

We achieved our goal of successfully replicating the univariate CSR model results!

