---
title: 2020 (Re)Insurer Stock Price Performance
author: ''
date: '2021-01-31'
slug: 2020-re-insurer-stock-price-performance
categories: []
tags:
  - insurance
  - stocks
draft: false
---

```{r setup, warning=FALSE,message=FALSE,include=FALSE}

library(knitr)

knitr::opts_chunk$set(
	fig.align = "center",
	message = FALSE,
	warning = FALSE,
	echo = FALSE,
	collapse = TRUE,
  comment = "#>",
  out.width = "120%"
)

library(tidyverse)
library(tidyquant)
library(popformatting)
library(highcharter)
library(flextable)
```

In this post, we review the 2020 stock price returns for US property and casualty (re)insurers. 

```{r import-stock-prices, message=FALSE, warning=FALSE, cache=TRUE}

start_stock <- as_date("2020-01-01")
end_stock <- as_date("2021-01-01")

tickers <- read_csv("E:/Projects/tickers.csv")

# get a list of all the Property-Casualty Insurers from the major US stock exchanges
stock_tickers <- list('AMEX','NASDAQ','NYSE') %>% map_df(tq_exchange) %>% 
  filter(!is.na(market.cap)) %>%
  mutate(market_cap = market.cap) %>%
  inner_join(tickers %>% select("symbol"=Ticker, "sector" = `Lockton Composite`), by = "symbol") %>% 
  dplyr::select(symbol, company, sector, market_cap) %>%
  arrange(desc(market_cap))

# list of mortgage / bond insurers to exclude
excluded_insurers <- c("MHLD", "CS", "KFS", "SCR", "OXBR")

stock_tickers <- stock_tickers %>%
  filter(!(symbol %in% excluded_insurers))

# now get the historical stock prices for all of the tickers; calculate daily returns
stock_prices <- stock_tickers %>% 
  tq_get(get = "stock.prices", from = start_stock, to = end_stock) %>% 
  filter(!is.na(adjusted)) %>% 
  dplyr::group_by(symbol) %>%
  tq_mutate(select = adjusted, mutate_fun = periodReturn, period = "daily", type = "arithmetic", 
            col_rename = "daily_returns") %>% 
  mutate(cum_return = adjusted / dplyr::first(adjusted) - 1)

# get S&P 500 historical prices; use for comparison
sp500_price <- tibble(symbol = "SPY") %>% tq_get(get = "stock.prices", 
                                                 from = start_stock, to = end_stock) %>%
  tq_mutate(select = adjusted, mutate_fun = periodReturn, period = "daily", type = "arithmetic", 
            col_rename = "daily_returns") %>% 
  mutate(cum_return = adjusted / dplyr::first(adjusted) - 1)


sp500_price <- sp500_price %>% mutate(sp500 = adjusted, sp500_returns = daily_returns)
stock_prices <- stock_prices %>% 
  left_join(sp500_price %>% dplyr::select(date, sp500, sp500_returns), by = "date")
```

```{r cum-change}
annual_returns <- stock_prices %>%
  group_by(symbol, company, sector) %>% 
  summarise(market_cap = market_cap[1], 
            boy_price = dplyr::first(close), eoy_price = dplyr::last(close),
            return = dplyr::last(cum_return), price_return = dplyr::last(close) / dplyr::first(close) - 1,
            div_return = (1 + return) / (1 + price_return) - 1,
            min_return = min(cum_return), max_return = max(cum_return),
            min_price = min(adjusted), max_price = max(adjusted), 
            avg_return = mean(daily_returns), sd_returns = sd(daily_returns),
            up_days = sum(daily_returns>0)/n(), down_days = sum(daily_returns<0)/n()) %>% 
  arrange(desc(return))
```

The top 5 performing stocks are shown in the graph below. All five of the stocks more than doubled the S&P500 return. KNSL and PLMR were both in the top 5 performing stocks in 2019. 

```{r highchart-top}
hc_data <- stock_prices %>% 
  filter(symbol %in% annual_returns$symbol[1:5]) %>%
  ungroup() %>% 
  mutate(cum_return = cum_return) %>% 
  select(date, symbol, cum_return) %>% 
  spread(key = symbol, value = cum_return) %>% 
  timetk::tk_xts()

hc_sp500 <- sp500_price %>% 
  select(date, symbol, cum_return) %>% 
  timetk::tk_xts()

hc <- highchart(type = "stock") %>% 
  hc_title(text = "Top 5 Performing P&C Stocks in 2020 - Annual Total Return") %>% 
  hc_subtitle(text = "KNSL, PLMR, ERIE, PGR, WTRE") %>% 
  hc_add_series(hc_data[,1], name = names(hc_data[,1])) %>% 
  hc_add_series(hc_data[,2], name = names(hc_data[,2])) %>%
  hc_add_series(hc_data[,3], name = names(hc_data[,3])) %>%
  hc_add_series(hc_data[,4], name = names(hc_data[,4])) %>%
  hc_add_series(hc_data[,5], name = names(hc_data[,5])) %>%
  hc_add_series(hc_sp500, name = "SPY") %>% 
  hc_tooltip(pointFormat = "<span style=\"color:{series.color}\">{series.name}</span>:<b>{point.y:.3f}</b><br/>",
             shared = TRUE) %>% 
  hc_legend(enabled = TRUE) %>% 
  hc_add_theme(hc_theme_smpl())

hc
```

If we focus on the largest five P&C stocks (by market cap). Only PGR outperformed the S&P500 with the other four stocks significantly underperforming. 

```{r highcharter-largest}
stock_prices$market_cap[stock_prices$symbol %in% c("ESGRO","ESGRP")] <- c(NA_real_, NA_real_)
annual_returns$market_cap[annual_returns$symbol %in% c("ESGRO","ESGRP")] <- c(NA_real_, NA_real_)

hc_data2 <- stock_prices %>% 
  filter(symbol %in% head(annual_returns %>% ungroup() %>% 
                            filter(!(sector %in% "L&H")) %>% arrange(desc(market_cap)) %>% pull(symbol),5)) %>%
  ungroup() %>% 
  mutate(cum_return = cum_return) %>% 
  select(date, symbol, cum_return) %>% 
  spread(key = symbol, value = cum_return) %>% 
  timetk::tk_xts()

hc2 <- highchart(type = "stock") %>% 
  hc_title(text = "2020 Stock Performance of 5 Largest P&C Stocks - Annual Total Return") %>% 
  hc_subtitle(text = "PGR, TRV, CB, ALL, AIG") %>% 
  hc_add_series(hc_data2[,1], name = names(hc_data2[,1])) %>% 
  hc_add_series(hc_data2[,2], name = names(hc_data2[,2])) %>%
  hc_add_series(hc_data2[,3], name = names(hc_data2[,3])) %>%
  hc_add_series(hc_data2[,4], name = names(hc_data2[,4])) %>%
  hc_add_series(hc_data2[,5], name = names(hc_data2[,5])) %>%
  hc_add_series(hc_sp500, name = "SPY") %>% 
  hc_tooltip(pointFormat = "<span style=\"color:{series.color}\">{series.name}</span>:<b>{point.y:.3f}</b><br/>",
             shared = TRUE) %>% 
  hc_legend(enabled = TRUE) %>% 
  hc_add_theme(hc_theme_smpl())

hc2
```

The table below summarizes the 2020 performance for all US listed P&C and Life (re)insurer stocks. 

```{r summary-table}
annual_returns %>%
  ungroup() %>% 
  transmute(ticker = symbol, Company = company, sector,
            `Total return` = popformatting::pct_format(return, 1),
            `Market cap ($m)` = popformatting::number_format(market_cap / 1e6, 0),
            `BOY price` = popformatting::number_format(boy_price, 2),
            `EOY price` = popformatting::number_format(eoy_price, 2),
            `Price return` = popformatting::pct_format(price_return, 1),
            `Div yield` = popformatting::pct_format(div_return, 1),
            `% up days` = popformatting::pct_format(up_days, 0),
            `% down days` = popformatting::pct_format(down_days, 0)
            ) %>% 
  flextable::regulartable() %>% 
  popformatting::ft_theme() %>% 
  align(j = 1:3, align = "left", part = "all") %>% 
  popformatting::add_title_header("Summary of 2020 (Re)Insurer Stock Performance")
```

